Towards a Cognitive System That Can Recognize Spatial Regions Based on Context

### 2012-01-25 code/scenarios/cdsr/trunk/papers/AAAI2012/main-aaai12-cdsr.tex
Kate: fixed typos
minor edits page 1

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a cognitive system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  We then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we implicitly represent context using the types of objects present in a room and their location relative to each other. The following sections describe how we construct these ingredients of context from the starting point of a robot that can sense its environment.

@@

\fw{
(regionType CDSR9 FrontRegion) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Room3)) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Group1))\\
...
}
}
  \caption{Representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains the symbolic description for the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}. The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region. \fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}. The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks. There are two more boundary segments completing a polygon for this region. The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room". This definition for the front of the room is specific to \fw{Room3} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were further to the left then the region would be smaller).

@@

Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones. Each oval represents a predicate, and the entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case . This is done in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Then SME will align the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}. While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the boundary segment expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping. The candidate inference is shown in red in the figure. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that results from this is then passed on to the QSR generator. In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area. To this we add the CDSR representing the front of the room. The case includes a total of 50 expression relating the 20 entities. Six of these expressions are used to define the boundary segments and CDSR representing the front of the room. The target case includes 26 expressions and 11 entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base. The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences. Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target. The green region in the lower image of Figure \ref{fig:rooms} illustrates the inferred CDSR.

>>>>

\maketitle
\begin{abstract}
In order to collaborate with people in the real world, cognitive systems must be able to represent and reason about spatial regions in human environments. Consider the command \emph{``go to the front of the classroom''}. The spatial region mentioned (the front of the classroom) is not perceivable using geometry alone. Instead it is defined by its functional use, implied by nearby objects and their configuration. In this paper, we define such areas as \textit{context-dependent spatial regions} and present a cognitive system able to learn them by combining qualitative spatial representations, semantic labels, and analogy. The system is capable of generating a collection of qualitative spatial representations describing the configuration of the entities it perceives in the world. It can then be taught context-dependent spatial regions using \textit{anchor points} defined on these representations.  From this we then demonstrate how an existing computational model of analogy can be used to detect context-dependent spatial regions in previously unseen rooms. To evaluate this process we compare detected regions to annotations made on maps of real rooms by human volunteers.
\end{abstract}

@@

The context which defines a CDSR is a combination of the functional and geometric properties of a room, i.e. what can be done there and where. In this work we implicitly represent context using the types of objects present in a room and their location relative to each other. The following sections describe how we construct symbolic representations of these ingredients of context from robot sensor data.

@@

\fw{(regionType CDSR9 FrontRegion) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMaxXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Room3)) \\
(boundarySegment CDSR9 \\
\hspace*{2em}(YMinXFewestFn Room3) \\
\hspace*{2em}(YMinXFewestFn Group1))
}
}
  \caption{Three of the five expressions representing the front of the classroom context-dependent region \fw{CDSR9}}
  \label{fig:cdsr-reps}
\end{figure}

Figure \ref{fig:cdsr-reps} contains three of the five expressions defining the front of classroom \fw{Room3} which is pictured in the top of Figure~\ref{fig:rooms}. The boundary segments (shown in orange in Figure~\ref{fig:rooms}) define the extent of the region. \fw{(YMaxXFewestFn Room3)} and \fw{(YMinXFewestFn Room3)} are the points with the highest and lowest y coordinate out of the set of points within 5\% of the minimum x coordinate of \fw{Room3}. The next segment connect the lower left coordinate in the figure to the \fw{(YMinXFewestFn Group1)}, where \fw{Group1} includes the eight desks. There are two more boundary segments completing a polygon for this region. The semantic label \fw{FrontRegion} ties this polygon to a conceptual region, "the front of the room". This definition for the front of the room is specific to \fw{Room3} and its entities. It is clearly context-dependent because its extent is dependent on the arrangement of the anchor points used to define its boundary. If the desks were in a different position then the region would cover a different extent (e.g. if they were further to the left then the region would be smaller).

@@

Figure \ref{fig:analogy} illustrates a sample mapping between six base expressions and three target ones. Each oval represents a predicate, and the entity arguments are represented by squares. SME generates a mapping between the base expressions \fw{(group Desk1 Desk2)} and  \fw{(xMinusYZero Desk1 Desk2)}, and the  target expressions \fw{(group Desk11 Desk12)} and \fw{(xMinusYZero Desk11 Desk12)} as well as between the \fw{regionType} expressions in each case in the following manner. First, the predicates of these expressions are placed in correspondence, as identical predicates are preferred by SME. Then SME aligns the arguments of the aligned predicates, \fw{Desk1} with \fw{Desk11},  \fw{Desk2} with \fw{Desk12}, and \fw{CDSR1} with \fw{CDSR2}. While there is another \fw{XMinusYZero} statement in the base about two desks, it cannot correspond to either of the target expressions in the same mapping due to the one-to-one constraint in SME which allows each element in the target to map to at most one element in the base and vice versa. In Figure \ref{fig:analogy}, the correspondences are highlighted by the hashed bi-directional arrows. Next, SME creates a candidate inference for the boundary segment expression, because both the mapped \fw{Group} and \fw{regionType} predicates participate in the mapping. The candidate inference is shown in red in the figure. Note that inference is selective, with no candidate inferences generated for the entirely unmapped expressions. %Next, we describe how we use the results of an analogy to infer a context-dependent spatial region in the target.

@@

We first create the base and target representations by running Dora in the two different classrooms. These are pictured in the top and bottom of Figure~\ref{fig:rooms} respectively. In each case, Dora is manually driven around the room to allow it to create a metric map. Once the map is created, Dora is then positioned such that the objects are observable and the visual recognition system is run. The map and object data that result from this are then passed on to the QSR generator. In the base case, Dora perceives 8 individual desks, a group entity containing these desks and the room area. To this we add the CDSR representing the front of the room. The case includes a total of 50 expression relating the 20 entities. Six of these expressions are used to define the boundary segments and CDSR representing the front of the room. The target case includes 26 expressions and 11 entities.

SME generates an analogy between the base and target cases enabling the transfer of the symbolic description of the front of the room to the new situation requiring \fw{Room3} and \fw{Group1} participate in the mapping as they are referenced by the anchor points in the base. The resulting analogy includes 26 correspondences between the entities and expressions and 32 candidate inferences. Four of these candidate inferences define the CDSR in the target with anchor points to the group of desks and the room in the target. The green region in the lower image of Figure \ref{fig:rooms} illustrates the transferred CDSR.
